\section{Literatur}

\begin{frame}[allowframebreaks]
  \frametitle{Literatur}    
  \begin{thebibliography}{KMT03}

    \bibitem[AB09]{AB09}
    Sanjeev Arora and Boaz Barak, \emph{Computational complexity: A modern
      approach}, 1st ed., Cambridge University Press, New York, NY, USA, 2009.

    \bibitem[Coo71]{C71}
    Stephen~A. Cook, \emph{The complexity of theorem-proving procedures},
      Proceedings of the third annual ACM symposium on Theory of computing (New
      York, NY, USA), STOC '71, ACM, 1971, pp.~151--158.

    \bibitem[CR79]{CR79}
    Stephen~A. Cook and Robert~A. Reckhow, \emph{The relative efficiency of
      propositional proof systems}, Journal of Symbolic Logic \textbf{44} (1979),
      36--50.

    \bibitem[For09]{F09}
    Lance Fortnow, \emph{The status of the {P} versus {NP} problem}, Commun. ACM
      \textbf{52} (2009), no.~9, 78--86.

    \bibitem[KM00]{KM00}
    Johannes K\"{o}bler and Jochen Messner, \emph{Is the standard proof system for
      {SAT} p-optimal?}, Proceedings of the 20th Conference on Foundations of
      Software Technology and Theoretical Computer Science (London, UK, UK), FST
      TCS 2000, Springer-Verlag, 2000, pp.~361--372.

    \bibitem[KMT03]{KMT03}
    Johannes K\"{o}bler, Jochen Messner, and Jacobo Tor\'{a}n, \emph{Optimal proof
      systems imply complete sets for promise classes}, Inf. Comput. \textbf{184}
      (2003), no.~1, 71--92.

    \bibitem[KP89]{KP89}
    Jan Kraj\'{\i}cek and Pavel Pudl{\'a}k, \emph{Propositional proof systems, the
      consistency of first order theories and the complexity of computations}, J.
      Symb. Log. \textbf{54} (1989), no.~3, 1063--1079.

    \bibitem[Mes99]{Mes99}
    Jochen Messner, \emph{On optimal algorithms and optimal proof systems},
      Proceedings of the 16th annual conference on Theoretical aspects of computer
      science (Berlin, Heidelberg), STACS'99, Springer-Verlag, 1999, pp.~541--550.

    \bibitem[Pap94]{Pap94}
    Christos~H. Papadimitriou, \emph{Computational complexity}, Addison-Wesley,
      1994.

  \end{thebibliography}

\end{frame}

\begin{frame}[allowframebreaks]
  \frametitle{Bildquellen}

  \begin{itemize}
   \item Kurt Gödel: Wikipedia, \texttt{\href{https://de.wikipedia.org/w/index.php?title=Datei:Kurt_g\%C3\%B6del.jpg\&filetimestamp=20110724092213}{de.wikipedia.org}}
  \end{itemize}  
\end{frame}
